---
title: "Donkey Games"
subtitle: "Final Project"
author: "Stuart McColl"
date: "January 2021"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}

# read in cleaned data and load libraries:

library(tidyverse)
library(rmarkdown)
library(pander)
library(latexpdf)
library(here)

here::here()

sales_all_clean <- read_csv(here("clean_data", "sales_all_clean"))
sales_xbox_ps4_clean <- read_csv(here("clean_data", "sales_xbox_ps4_clean"))
shipped_2019_clean <- read_csv(here("clean_data", "shipped_2019_clean"))

```


# Introduction 

The following report provides an exploratory analysis into the sales of video games. There are three data sets under observation within this report: 

The first data set contains the sales of video games across all the platforms and includes a wide range of games from 1970 up to 2019. The data set contains sales recorded in 2016 and 2019 respectively.

The second data set contains the details of all the shipped games from 2019 and also includes a wide range of games released in the early eighties up to 2019.

The final data set contains the sales of XBox One and PS4 games released from 2013 to 2019.

# Game Sales 2016 and 2019 

### General Observations

**A brief glimpse at the data:**

```{r, echo = FALSE}

sales_all_clean %>% 
  head()

```

<br>

**What are the top 10 games in 2016?**

```{r,echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(sales_year == 2016) %>% 
  arrange(desc(global_sales)) %>% 
  slice_max(global_sales, n = 10)

```
The results are interesting; in 2016 **Nintendo** fill the entire top ten with a mix of retro games and games for the Wii console. Super Mario Bros, which was originally released in 1985 ranked number two in the best selling games. Taking first place is Wii Sports.

<hr>

**What are the top 10 games in 2019?**

```{r,echo = FALSE, include = T}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(sales_year == 2019) %>% 
  arrange(desc(global_sales)) %>% 
  slice_max(global_sales, n = 10)

```

<br>

How do they compare?

```{r, echo = FALSE}

tibble(
  "2016" = c("Wii Sports (Wii)", "Super Mario Bros. (NES)", 
             "Mario Kart Wii (Wii)", "Wii Sports Resort (Wii)",
             "Pokemon Red/Pokemon Blue (GB)", "Tetris (GB)",
             "New Super Mario Bros. (DS)", "Wii Play (Wii)",
             "New Super Mario Bros. Wii (Wii)", "Duck Hunt (NES)"),
  
  "2019" = c("Grand Theft Auto V (PS3)", "Grand Theft Auto V (PS4)",
             "Grand Theft Auto: Vice City (PS2)", "Grand Theft Auto V (X360)",
             "Call of Duty: Black Ops 3 (PS4)", 
             "Call of Duty: Modern Warfare 3 (X360)", 
             "Call of Duty: Black Ops (X360)",
             "Red Dead Redemption 2 (PS4)", "Call of Duty: Black Ops II (X360)",
             "Call of Duty: Black Ops II (PS3)")
) %>% 
  pander()

```
Interestingly, **Nintendo** have been knocked out the top ten in 2019. **Grand Theft Auto** and **Call of Duty** being the top ranked games.

<hr>

**What are the top game genres of 2016?**

```{r, message=FALSE, echo = FALSE}

sales_all_clean %>% 
  filter(sales_year == 2016) %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "in 2016",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750),
                     expand = c(0.025, 0.025))


```

<br>

**What are the top game genres of 2019?**

```{r, message=FALSE, echo = FALSE}

sales_all_clean %>% 
  filter(sales_year == 2019) %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "in 2019",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500),
                     expand = c(0.025, 0.025))


```

<br>

How do they compare?

```{r, echo = FALSE}

tibble(
  "2016" = c("Action", "Sports", "Shooter", "Role-Playing", "Platform", "Other",
             "Racing", "Fighting", "Simulation", "Puzzle", "Adventure", 
             "Strategy"),
  "2019" = c("Action", "Sports", "Shooter", "Other", "Racing", "Role-Playing",
             "Platform", "Fighting", "Adventure", "Simulation", "Strategy",
             "Puzzle")
) %>% 
  pander()

```
Interesting results; you can see that **Action**, **Sports** and **Shooter** 
games maintained their position as top 3 games genres. **Racing** games have 
made their way into the top 5.

### Action Games

```{r, echo = FALSE, include = F}

# 2016

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Action") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE, include = F}

# 2019

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Action") %>% 
  filter(sales_year == 2019) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE}

tibble(
  "2016" = c("Grand Theft Auto V (PS3)", "Grand Theft Auto: San Andreas (PS2)",
             "Grand Theft Auto V (X360)", "Grand Theft Auto: Vice City (PS2)",
             "Grand Theft Auto III (PS2)"),
  
  "2019" = c("Grand Theft Auto V (PS3)", "Grand Theft Auto V (PS4)",
             "Grand Theft Auto: Vice City (PS2)", "Grand Theft Auto V (X360)",
             "Red Dead Redemption 2 (PS4)")
) %>% 
  pander()

```
**Grand Theft Auto** clearly dominates the **Action** category.

<hr>

### Sports Games

```{r, echo = FALSE, include = F}

# 2016

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Sports") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE, include = F}

# 2019

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Sports") %>% 
  filter(sales_year == 2019) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE}

tibble(
  "2016" = c("Wii Sports (Wii)", "Wii Sports Resort (Wii)", "Wii Fit (Wii)",
             "Wii Fit Plus (Wii)", "FIFA 16 (PS4)"),
  
  "2019" = c("FIFA 18 (PS4)", "FIFA 17 (PS4)", "FIFA 19 (PS4)", "FIFA 16 (PS4)",
             "FIFA 13 (PS3)")
) %>% 
  pander()

```
**Wii Sports** clearly wins this category. **FIFA** also highly ranked in this 
category.

<hr>

### Shooter Games

```{r, echo = FALSE, include = F}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Shooter") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE, include = F}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Shooter") %>% 
  filter(sales_year == 2019) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE}

tibble(
  "2016" = c("Duck Hunt (NES)", "Call of Duty: Modern Warfare 3 (X360)",
             "Call of Duty: Black Ops 3 (PS4)", "Call of Duty: Black Ops (X360)",
             "Call of Duty: Black Ops II (PS3)"),
  
  "2019" = c("Call of Duty: Black Ops 3 (PS4)", 
             "Call of Duty: Modern Warfare 3 (X360)", 
             "Call of Duty: Black Ops (X360)",
             "Call of Duty: Black Ops II (X360)",
             "Call of Duty: Black Ops II (PS3)"
             )
) %>% 
  pander()

```
**Duck Hunt**, the classic Nintendo game from 1984 tops this category. **Call of Duty** tops the 'modern' games in this category.

<hr>

### Role Playing Games

```{r, echo = FALSE, include = T}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Role-Playing") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 5)

```

```{r, echo = FALSE, include = T}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Role-Playing") %>% 
  filter(sales_year == 2019) %>% 
  slice_max(global_sales, n = 5)

```

```{r}

tibble(
  "2016" = c("Pokemon Red/Pokemon Blue (GB)", "Pokemon Gold/Pokemon Silver (GB)",
             "Pokemon Diamond/Pokemon Pearl (DS)",
             "Pokemon Ruby/Pokemon Sapphire (GBA)",
             "Pokemon Black/Pokemon White (DS)")
) %>% 
  pander()

```



**Pokemon** are the clear winners from this category.

<hr>

### Platform Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Platform") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```

**Super Mario** wins.

<hr>

### Other Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Other") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```
This is an interesting category as there is quite a mix of top games. It is clear that the Wii is popular, especially with interactive games, such as **Wii Play** and **Just Dance** scoring highly.

<hr>

### Racing Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Racing") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```

**Mario Kart** taking lead in this category. Special mention to **Gran Turismo** also as the top 'non-Nintendo' game.

<hr>

### Fighting Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Fighting") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```

**Super Smash Bros. Brawl** for the Wii taking first place in this category. Classic fighting games such as **Tekken** and **Street Fighter** are very popular.

<hr>

### Simulation Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Simulation") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```
Interesting result from this category, **Nintendogs** is top in global sales. **The Sims** and **Animal Crossing** also performing well.

<hr>

### Puzzle Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Puzzle") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```
**Tetris** is the top classic puzzle game. **Brain Age** and **Professor Layton** games performed very well in this category.

<hr>

### Adventure Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Adventure") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```
**Super Mario Land 2: 6 Golden Coins** is the number one 'Adventure' game, released in 1992 for the GameBoy. **Assassin's Creed** is the top 'modern' game in this category.

<hr>

### Strategy Games

```{r, echo = FALSE}

sales_all_clean %>% 
  select(name, console, year_of_release, genre, publisher, global_sales, sales_year) %>% 
  filter(genre == "Strategy") %>% 
  filter(sales_year == 2016) %>% 
  slice_max(global_sales, n = 20)

```
**Pokemon Stadium** is the top game in this category.

<hr>




