---
title: "Donkey Games"
subtitle: "Final Project"
author: "Stuart McColl"
date: "January 2021"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}

# read in cleaned data and load libraries:

library(tidyverse)
library(rmarkdown)
library(pander)
library(latexpdf)
library(here)

here::here()

clean_sales_data <- read_csv(here("clean_data", "clean_sales_data.csv"))


```


# Introduction 

The following report provides an exploratory analysis into the sales of video games. The data used for this report contains the global sales of games dating as far back as 1970 through to the present. Having such a vast range of games sales allows for a interesting look in to trends in top games across the years in order to identify what types of games, historically have performed well.

# The Data

### General Observations

**A brief glimpse at the data:**

```{r, echo = FALSE}

clean_sales_data %>% 
  head()

```
The data is an amalgamation of 4 data sets, each with their own set of sales and useful information. The final data contains the most up to date sales figures for each game as of 2019.

<br>

# Games History

### 1970's

```{r, echo = FALSE, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "1970" & year_of_release <= "1979") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) 
  
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1970" & year_of_release <= "1979") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1970s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5),
                     expand = c(0.025, 0.025))

ggsave("1970_plot.png", path = here("plots"))
  
  

```

<br>

### 1980's

```{r, echo = FALSE, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "1980" & year_of_release <= "1989") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1980" & year_of_release <= "1989") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1980s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130),
                     expand = c(0.025, 0.025))


ggsave("1980_plot.png", path = here("plots"))

```

<br>

### 1990's

```{r, echo = FALSE, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "1990" & year_of_release <= "1999") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1990" & year_of_release <= "1999") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1990s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200),
                     expand = c(0.025, 0.025))

ggsave("1990_plot.png", path = here("plots"))

```

<br>

### 2000's

```{r, echo = FALSE, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "2000" & year_of_release <= "2010") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "2000" & year_of_release <= "2010") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "2000's",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 200, 400, 600, 800, 1000, 1200, 1400),
                     expand = c(0.025, 0.025))

ggsave("2000_plot.png", path = here("plots"))

```

<br>

### 2010's

```{r, echo = FALSE, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "2010" & year_of_release <= "2020") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "2010" & year_of_release <= "2020") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "2010's",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 200, 400, 600, 800, 1000, 1200, 1400),
                     expand = c(0.025, 0.025))

ggsave("2010_plot.png", path = here("plots"))

```

<br>

# Games Genres

### Action and Adventure Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Action-Adventure") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Grand Theft Auto** clearly dominates the **Action-Adventure** category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Action-Adventure") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Action and Adventure Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000),
                     expand = c(0.025, 0.025))

ggsave("action_adventure_plot.png", path = here("plots"))



```

<hr>

### Sports Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Sports") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Wii Sports** clearly wins this category. **FIFA** also highly ranked in this 
category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Sports") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Sports Games", 
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800),
                     expand = c(0.025, 0.025))

ggsave("sports_plot.png", path = here("plots"))



```

<hr>

### Shooter Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Shooter") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Call of Duty** tops the **shooter** games in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Shooter") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Shooter Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600),
                     expand = c(0.025, 0.025))

ggsave("shooter_plot.png", path = here("plots"))



```

<hr>

### Role Playing Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Role-Playing") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Pokemon** are the clear winners from this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Role-Playing") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Role Playing Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400),
                     expand = c(0.025, 0.025))



```

<hr>

### Platform Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Platform") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Super Mario** clearly dominates the **platform** games category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Platform") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Platform Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400),
                     expand = c(0.025, 0.025))



```

<hr>

### Other Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Other") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
This is an interesting category as there is quite a mix of top games. **Minecraft** takes the top spot. **Just Dance** also performing well i nthis category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Other") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Other Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500),
                     expand = c(0.025, 0.025))



```

<hr>

### Racing Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Racing") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Mario Kart** taking lead in this category. **Gran Turismo** also performing very well.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Racing") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Racing Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500),
                     expand = c(0.025, 0.025))



```

<hr>

### Fighting Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Fighting") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Super Smash Bros. Brawl** taking first place in this category. Classic fighting games such as **Tekken**, **Mortal Kombat** and **Street Fighter** are very popular.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Fighting") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Fighting Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250),
                     expand = c(0.025, 0.025))



```

<hr>

### Simulation Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Simulation") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
Interesting result from this category, **Nintendogs** is top in global sales. **The Sims** and **Animal Crossing** also performing well.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Simulation") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Simulation Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250),
                     expand = c(0.025, 0.025))



```

<hr>

### Puzzle Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Puzzle") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```
**Tetris** is the top classic puzzle game. **Brain Age** performed very well in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Puzzle") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Puzzle Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100, 120),
                     expand = c(0.025, 0.025))



```

<hr>

### Strategy Games

```{r, echo = FALSE, include = T, message=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Strategy") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```

**Pokemon Stadium** is the top game in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Strategy") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Strategy Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000),
                     expand = c(0.025, 0.025))



```

<hr>

# Regional Sales 

### All Regions

```{r, include=F}

clean_sales_data %>% 
  summarise(north_america_sales = sum(na_sales)/sum(global_sales)*100, 
            japan_sales = sum(jp_sales)/sum(global_sales)*100,
            europe_sales = sum(eu_sales)/sum(global_sales)*100,
            other_sales = sum(other_sales)/sum(global_sales)*100) 

```


```{r, echo=FALSE, message=FALSE}

tibble(
  `North America` = 0.49,
  `Japan` = 0.14,
  `Europe` = 0.28,
  `Rest of the World` = 0.09
) %>% 
  gather(key = Region, value = Percent) %>% 
  
  ggplot(aes(x = reorder(Region, -Percent), y = Percent)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  theme_light() +
  labs(title = "Sales by Region", 
       subtitle = "from 1970 to 2020",
       x = " ", 
       y = " ") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) 

ggsave("region_plot.png", path = here("plots"))


```

### North America

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(na_sales_millions = sum(na_sales)) %>% 
  arrange(desc(na_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, na_sales_millions), 
             y = na_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "North America Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000, 2250),
                     expand = c(0.025, 0.025))

ggsave("north_america_plot.png", path = here("plots"))


```

### Europe

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(eu_sales_millions = sum(eu_sales)) %>% 
  arrange(desc(eu_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, eu_sales_millions), 
             y = eu_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Europe Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 1750),
                     expand = c(0.025, 0.025))

ggsave("europe_plot.png", path = here("plots"))


```

### Japan

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(jp_sales_millions = sum(jp_sales)) %>% 
  arrange(desc(jp_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, jp_sales_millions), 
             y = jp_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Japan Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300, 350, 400),
                     expand = c(0.025, 0.025))

ggsave("japan_plot.png", path = here("plots"))


```

### Rest of the World

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(other_sales_millions = sum(other_sales)) %>% 
  arrange(desc(other_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, other_sales_millions), 
             y = other_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Rest of the World Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225),
                     expand = c(0.025, 0.025))

ggsave("rest_of_world_plot.png", path = here("plots"))


```

# Top Games

### Top 10 Games

```{r, include=T, message=FALSE, echo=FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales)) %>% 
  slice_max(total_global_sales, n = 10)

```

```{r, warning=FALSE, message=FALSE, include=FALSE}

# North America Best Game
clean_sales_data %>% 
  group_by(name) %>% 
  select(-c(eu_sales, jp_sales, global_sales, other_sales)) %>% 
  summarise(total_na_sales = sum(na_sales)) %>% 
  arrange(desc(total_na_sales)) 

# Europe Best Game
clean_sales_data %>% 
  group_by(name) %>% 
  select(-c(na_sales, jp_sales, global_sales, other_sales)) %>% 
  summarise(total_eu_sales = sum(eu_sales)) %>% 
  arrange(desc(total_eu_sales)) 

# Japan Best Game
clean_sales_data %>% 
  group_by(name) %>% 
  select(-c(eu_sales, na_sales, global_sales, other_sales)) %>% 
  summarise(total_jp_sales = sum(jp_sales)) %>% 
  arrange(desc(total_jp_sales)) 


```


The results are interesting; **Nintendo** taking up 60% of the top ten with the top game being Wii Sports.

<hr>

### Top Genres

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000, 2250),
                     expand = c(0.025, 0.025))

ggsave("top_genres_plot.png", path = here("plots"))


```

# Recommendations

Well it's not that straight forward. The top performing games, especially the likes of Wii Sports are pretty unique and could have easily gone the other way and not been a hit at all. 

I think it's pretty safe to say that Action and Adventure games do well. Grand Theft Auto is the best example. People love it for many reasons, mostly:

1. Aesthetics - the graphics are very good and it's great to look at.

2. Characters - you are able to choose from different protagonists, who are unique.

3. It's also a very good multi-player game.

Sports games may be a little risky to be honest. Wii Sports is a bit of an outlier and FIFA pretty much dominates the football category. 

Shooter games could be a recommendation from me. In terms of trends, sales have never gone down. I think fun, nostalgia games such as Duck Hunt or even the N64 classic Golden Eye, which is an absolute cult classic could be a suitable route. However, this should perhaps be left to Nintendo and actually producing something more in line with Call of Duty might be more appropriate. 

It also depends on what market you intend to launch the game. North America and Europe are fairly similar in terms to the top performing games, however if you were to only target Japan. Action and Adventure, Sports and Shooter games would not be the best type of game as they are more into Role Playing games and in particular, Pokemon!

It's important to learn what games sell many copies and to understand why they are popular, however I think ultimately, the most important thing for creating a new game is that it needs to be unique and creative in order to stand out. 

<hr>













