---
title: "Donkey Games"
subtitle: "Final Project"
author: "Stuart McColl"
date: "January 2021"
output:
  html_document:
    df_print: paged
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, include = FALSE}

# read in cleaned data and load libraries:

library(tidyverse)
library(rmarkdown)
library(pander)
library(latexpdf)
library(here)

here::here()

clean_sales_data <- read_csv(here("clean_data", "clean_sales_data.csv"))


```


# Introduction 

The following report provides an exploratory analysis into the sales of video games. The data used for this report contains the global sales of games dating as far back as 1970 through to the present. Having such a vast range of games sales allows for a interesting look in to trends in top games across the years in order to identify what types of games, historically have performed well.

# The Data

### General Observations

**A brief glimpse at the data:**

```{r, echo = FALSE}

clean_sales_data %>% 
  head()

```
The data is an amalgamation of 4 data sets, each with their own set of sales and useful information. The final data contains the most up to date sales figures for each game as of 2019.

<br>

# Games History

### 1970's

```{r, echo = FALSE}

clean_sales_data %>% 
  filter(year_of_release >= "1970" & year_of_release <= "1979") 
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1970" & year_of_release <= "1979") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1970s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5),
                     expand = c(0.025, 0.025))

ggsave("1970_plot.png", path = here("plots"))
  
  

```

<br>

### 1980's

```{r, echo = FALSE}

clean_sales_data %>% 
  filter(year_of_release >= "1980" & year_of_release <= "1989") 
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1980" & year_of_release <= "1989") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1980s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130),
                     expand = c(0.025, 0.025))


ggsave("1980_plot.png", path = here("plots"))

```

### 1990's

```{r, echo = FALSE}

clean_sales_data %>% 
  filter(year_of_release >= "1990" & year_of_release <= "1999") 
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "1990" & year_of_release <= "1999") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "1990s",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200),
                     expand = c(0.025, 0.025))

ggsave("1990_plot.png", path = here("plots"))

```

### 2000's

```{r, echo = FALSE}

clean_sales_data %>% 
  filter(year_of_release >= "2000" & year_of_release <= "2010") 
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "2000" & year_of_release <= "2010") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "2000's",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 200, 400, 600, 800, 1000, 1200, 1400),
                     expand = c(0.025, 0.025))

ggsave("2000_plot.png", path = here("plots"))

```

### 2010's

```{r, echo = FALSE}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(year_of_release >= "2010" & year_of_release <= "2020") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))
 
```

```{r, echo = FALSE, message = F}

clean_sales_data %>% 
  filter(year_of_release >= "2010" & year_of_release <= "2020") %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "2010's",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 200, 400, 600, 800, 1000, 1200, 1400),
                     expand = c(0.025, 0.025))

ggsave("2010_plot.png", path = here("plots"))

```

# Games Genres

### Action and Adventure Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  group_by(name) %>% 
  filter(genre == "Action-Adventure") %>% 
  summarise(total_global_sales = sum(global_sales)) %>% 
  arrange(desc(total_global_sales))

```
**Grand Theft Auto** clearly dominates the **Action-Adventure** category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Action-Adventure") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Action and Adventure Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000),
                     expand = c(0.025, 0.025))

ggsave("action_adventure_plot.png", path = here("plots"))



```

<hr>

### Sports Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Sports") %>% 
  slice_max(global_sales, n = 5)

```
**Wii Sports** clearly wins this category. **FIFA** also highly ranked in this 
category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Sports") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Sports Games", 
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800),
                     expand = c(0.025, 0.025))

ggsave("sports_plot.png", path = here("plots"))



```

<hr>

### Shooter Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Shooter") %>% 
  slice_max(global_sales, n = 5)

```
**Duck Hunt**, the classic Nintendo game from 1984 tops this category. **Call of Duty** tops the 'modern' games in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Shooter") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Shooter Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600),
                     expand = c(0.025, 0.025))

ggsave("shooter_plot.png", path = here("plots"))



```

<hr>

### Role Playing Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Role-Playing") %>% 
  slice_max(global_sales, n = 5)

```
**Pokemon** are the clear winners from this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Role-Playing") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Role Playing Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400),
                     expand = c(0.025, 0.025))



```

<hr>

### Platform Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Platform") %>% 
  slice_max(global_sales, n = 5)

```
**Super Mario** wins.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Platform") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Platform Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400),
                     expand = c(0.025, 0.025))



```

<hr>

### Other Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Other") %>% 
  slice_max(global_sales, n = 5)

```
This is an interesting category as there is quite a mix of top games. It is clear that the Wii is popular, especially with interactive games, such as **Wii Play** and **Just Dance** scoring highly.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Other") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Other Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500),
                     expand = c(0.025, 0.025))



```

<hr>

### Racing Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Racing") %>% 
  slice_max(global_sales, n = 5)

```
**Mario Kart** taking lead in this category. Special mention to **Gran Turismo** also as the top 'non-Nintendo' game.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Racing") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Racing Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500),
                     expand = c(0.025, 0.025))



```

<hr>

### Fighting Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Fighting") %>% 
  slice_max(global_sales, n = 5)

```
**Super Smash Bros. Brawl** for the Wii taking first place in this category. Classic fighting games such as **Tekken** and **Street Fighter** are very popular.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Fighting") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Fighting Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250),
                     expand = c(0.025, 0.025))



```

<hr>

### Simulation Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Simulation") %>% 
  slice_max(global_sales, n = 5)

```
Interesting result from this category, **Nintendogs** is top in global sales. **The Sims** and **Animal Crossing** also performing well.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Simulation") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Simulation Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250),
                     expand = c(0.025, 0.025))



```

<hr>

### Puzzle Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Puzzle") %>% 
  slice_max(global_sales, n = 5)

```
**Tetris** is the top classic puzzle game. **Brain Age** performed very well in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Puzzle") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Puzzle Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 20, 40, 60, 80, 100, 120),
                     expand = c(0.025, 0.025))



```

<hr>

### Strategy Games

```{r, echo = FALSE, include = T}

clean_sales_data %>% 
  filter(genre == "Strategy") %>% 
  slice_max(global_sales, n = 5)

```

**Pokemon Stadium** is the top game in this category.

```{r, echo = FALSE, message = F, warning = FALSE}

clean_sales_data %>% 
  filter(genre == "Strategy") %>% 
  mutate(year_of_release = as.numeric(year_of_release)) %>% 
  filter(year_of_release != "Unknown") %>% 
  mutate(decade = floor(year_of_release / 10) * 10) %>% 
  filter(decade != 2020) %>% 
  group_by(decade) %>% 
  summarise(decade_sales = sum(global_sales)) %>% 
  
  ggplot(aes(x = decade, y = decade_sales)) +
  geom_area(alpha = 0.5, fill = "#666633") +
  geom_line(colour = "#cc9900", size = 1.2) +
  geom_point(colour = "#666633") +
  labs(title = "Strategy Games",
       x = "Decade", 
       y = "Global Sales (millions)") +
  theme_light() +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 700, 800, 900, 1000),
                     expand = c(0.025, 0.025))



```

<hr>

# Regional Sales 

### All Regions

```{r, include=F}

clean_sales_data %>% 
  summarise(north_america_sales = sum(na_sales)/sum(global_sales)*100, 
            japan_sales = sum(jp_sales)/sum(global_sales)*100,
            europe_sales = sum(eu_sales)/sum(global_sales)*100,
            other_sales = sum(other_sales)/sum(global_sales)*100) 

```


```{r, echo=FALSE, message=FALSE}

tibble(
  `North America` = 0.49,
  `Japan` = 0.14,
  `Europe` = 0.28,
  `Rest of the World` = 0.09
) %>% 
  gather(key = Region, value = Percent) %>% 
  
  ggplot(aes(x = reorder(Region, -Percent), y = Percent)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  theme_light() +
  labs(title = "Sales by Region", 
       subtitle = "from 1970 to 2020",
       x = " ", 
       y = " ") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) 

ggsave("region_plot.png", path = here("plots"))


```

### North America

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(na_sales_millions = sum(na_sales)) %>% 
  arrange(desc(na_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, na_sales_millions), 
             y = na_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "North America Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000, 2250),
                     expand = c(0.025, 0.025))

ggsave("north_america_plot.png", path = here("plots"))


```

### Europe

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(eu_sales_millions = sum(eu_sales)) %>% 
  arrange(desc(eu_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, eu_sales_millions), 
             y = eu_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Europe Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 100, 200, 300, 400, 500, 600, 1750),
                     expand = c(0.025, 0.025))

ggsave("europe_plot.png", path = here("plots"))


```

### Japan

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(jp_sales_millions = sum(jp_sales)) %>% 
  arrange(desc(jp_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, jp_sales_millions), 
             y = jp_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Japan Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 50, 100, 150, 200, 250, 300, 350, 400),
                     expand = c(0.025, 0.025))

ggsave("japan_plot.png", path = here("plots"))


```

### Rest of the World

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(other_sales_millions = sum(other_sales)) %>% 
  arrange(desc(other_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, other_sales_millions), 
             y = other_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Rest of the World Sales", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Sales (millions)") +
   scale_y_continuous(breaks = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225),
                     expand = c(0.025, 0.025))

ggsave("rest_of_world_plot.png", path = here("plots"))


```

# Top Games

### Top 10 Games

```{r,echo = FALSE}

clean_sales_data %>% 
  slice_max(global_sales, n = 10)

```
The results are interesting; **Nintendo** fill the entire top ten with a mix of retro games and games for the Wii console. Super Mario Bros, which was originally released in 1985 ranked number two in the best selling games. Taking first place is Wii Sports.

<hr>

### Top Genres

```{r, message=FALSE, echo = FALSE}

clean_sales_data %>% 
  group_by(genre) %>% 
  summarise(global_sales_millions = sum(global_sales)) %>% 
  arrange(desc(global_sales_millions)) %>% 
  
  ggplot(aes(x = reorder(genre, global_sales_millions), 
             y = global_sales_millions)) +
  geom_col(alpha = 0.8, colour = "white", fill = "#cc9900") +
  coord_flip() +
  theme_light() +
  labs(title = "Top Performing Games Genres", 
       subtitle = "from 1970 to 2020",
       x = "Genre", 
       y = "Global Sales (millions)") +
   scale_y_continuous(breaks = c(0, 250, 500, 750, 1000, 1250, 1500, 1750, 2000, 2250),
                     expand = c(0.025, 0.025))

ggsave("top_genres_plot.png", path = here("plots"))


```

# Recommendations













